---
title: "SA1 2.5 and 3"
author: "Zyann Lynn Mayo"
date: "2025-03-16"
output: html_document
runtime: shiny
---
## **2. Bivariate**

```{r}
library(shiny)

ui <- fluidPage(
    titlePanel("Bivariate Distribution Input"),
    sidebarLayout(
        sidebarPanel(
            textInput("xvals", "Enter X values (comma-separated)", "0,1"),
            textInput("yvals", "Enter Y values (comma-separated)", "0,1,2"),
            textAreaInput("probs", "Enter joint probabilities (row-wise)", "0.1,0.2,0.1\n0.3,0.2,0.1"),
            actionButton("compute", "Compute")
        ),
        mainPanel(
            tableOutput("marginals"),
            plotOutput("pdf_plot")
        )
    ),
    tags$style(HTML("
        body { background-color: #d97ee6; } 
        h1, h2 { color: #29042e; } 
        .btn { background-color: #291b2b; color: white; } 
    "))
)

server <- function(input, output) {
    observeEvent(input$compute, {
      X <- as.numeric(unlist(strsplit(input$xvals, ",")))
      Y <- as.numeric(unlist(strsplit(input$yvals, ",")))
  
      # Fix: Properly clean and split probabilities
      probs_vector <- as.numeric(unlist(strsplit(gsub("\n", ",", input$probs), ",")))
  
      # Debugging: Print lengths to check
      print(paste("X:", length(X), "Y:", length(Y), "Probs:", length(probs_vector)))
  
      # Ensure correct dimensions: rows = length(X), columns = length(Y)
      expected_length <- length(X) * length(Y)
  
      if (length(probs_vector) != expected_length) {
          showNotification(
              paste("Error: Probability matrix should have", expected_length, 
                    "values, but got", length(probs_vector)),
              type = "error"
          )
          return()
      }
  
      probs <- matrix(probs_vector, nrow = length(X), byrow = TRUE)
  
      # Compute marginal distributions
      P_X <- rowSums(probs)
      P_Y <- colSums(probs)
  
      output$marginals <- renderTable(data.frame(X_Prob = P_X, Y_Prob = P_Y))
      output$pdf_plot <- renderPlot(barplot(P_X, names.arg = X, col = "pink"))
  })


}

shinyApp(ui, server)

```

## **3. Markov Property**

Based on the given, let us first generate 10,000 searches going through a list of sites until the key phrase is found.
```{r}
set.seed(920)
searches <- 10000
p <- 0.6 
simulation <- rgeom(searches, p) + 1 #generated values should start from 1

hist(simulation, probability = TRUE, breaks = 30, 
     main = "Simulated Probability Density Function", 
     xlab = "Number of Searches Until Success", 
     col = "purple", border = "black")
```

Then, we compute the mean and variance:
```{r}
mean_sim <- mean(simulation)
var_sim <- var(simulation)
```

Which gives us the values **`r round(mean_sim, 2)`** for the **mean**, and **`r round(var_sim, 2)`** for the **variance**.

### **Obtain the simulated conditional distribution of searches where three searches have been carried out without success.**

```{r}
cond_sim <- simulation[simulation > 3]
mean_cond <- mean(cond_sim)
var_cond <- var(cond_sim)
```

After the condition, we get the values: **Mean =`r round(mean_cond, 2)`**, **Variance = `r round(var_cond, 2)`**.

### **Markov memoryless property**
According to the markov property of the geometric distribution, if X is a geometric random variable, then for each integer n:

\[
P(X = x + n|X > n) = P(X = x)

\]
**(a) P(X = 4|X >3) and P(X = 1)**
```{r}
p_x4_given_x3 <- sum(simulation == 4) / sum(simulation > 3)
p_x1 <- mean(simulation == 1)
```

These values show that the Markov property hold as the values are **approximately** equal to each other.

P(X = 4|X >3) = **`r round(p_x4_given_x3, 2)`**

P(X = 1) = **`r round(p_x1, 2)`**



**(b) P(X = 5|X >3) and P(X = 2)**
```{r}
p_x5_given_x3 <- sum(simulation == 5) / sum(simulation > 3)
p_x2 <- mean(simulation == 2)
```

These values show that the Markov property hold as the values are **approximately** equal to each other.

P(X = 4|X >3) = **`r round(p_x5_given_x3, 2)`** 

P(X = 1) = **`r round(p_x2, 2)`**






