---
title: "Rolling a Die Relative Frequency"
date: "2025-02-05"
output: html_document
---

Instructions: An experiment consists of rolling a die. Use R to simulate this experiment 600 times and obtain the relative frequency of each possible outcome. Hence, estimate the probability of getting each of 1, 2, 3, 4, 5, and 6.

First, simulate rolling a die 600 times and store it in the variable "rolls".

```{r}
rolls <- sample(1:6, size = 600, replace = TRUE)
```

Here are the first ten simulations:

```{r}
head(rolls,10)
```

Let's summarize how often each possible outcome appeared in the 600 rolls.

```{r}
table(rolls)
```

Finally, let's calculate the relative frequency or probability of getting each possible outcome in rolling a die.
```{r, fig.align='center'}
dice_prob <- table(rolls) / 600
dice_prob

barplot(dice_prob)
```

These values estimate the probability of each outcome of rolling a fair die which should ideally be: $$\frac{1}{6} \approx 0.1667$$.

