---
title: "NO2 FA3"
author: "Zyann Lynn Mayo"
date: "2025-02-16"
output: pdf_document
header-includes:
  - \usepackage{changepage}
---


## **2. Binary Communication Channel**

\begin{adjustwidth}{1cm}{1cm} 
Instructions: A binary communication channel carries data as one of two sets of signals denoted by 0 and 1. Owing to noise, a transmitted 0 is sometimes received as a 1, and a transmitted 1 is sometimes received as a 0. For a given channel, it can be assumed that a transmitted 0 is correctly received with probability 0.95, and a transmitted 1 is correctly received with probability 0.75. Also, 70% of all messages are transmitted as a 0. 
\end{adjustwidth}

Let's insert the given into variables

```{r}
# ZERO

# probability of correctly received 0
prob_c_rzero <- 0.95
# probability of incorrectly received (1 was received)
prob_w_rzero <- (1-0.95)
# probability of 0 transmitted
prob_tzero <- 0.70

# ONE

# probability of correctly received 1
prob_c_rone <- 0.75
# probability of incorrectly received (0 was received)
prob_w_rone <- (1-0.75)
# probability of 1 transmitted
prob_tone <- (1-0.70)

```

### **(a) a 1 was received**

The formula of the probablity of the event that 1 was received is the union of receiving 1 from either 0 and 1 that was transmitted.

\[
P(R_1) = (T_1 \cap R_1) \cup (T_0 \cap R_1)
\]

which is also:

\[
P(R_1) = P(T_1) P(R_1|T_1) + P(T_0) P(R_1|T_0) 
\]

Substituting the values...

```{r}
prob_rone <- (prob_tone*prob_c_rone) + (prob_tzero*prob_w_rzero)
#          = (0.3 * 0.75) + (0.7 * 0.05)
```

We get that the probability of receiving 1 is:
\[
P(R_1) = `r prob_rone`
\]

### **(b) a 1 was transmitted given that a 1 was received**

Using Bayes' rule for two events, the formula to be used will be:

\[
P(T_1|R_1) = \frac{P(T_1) P(R_1|T_1)}{P(R_1)}
\]

Using the initial given values and the value of that was acquired in the previous item...

```{r}
prob_tone_rone <- (prob_tone*prob_c_rone) / prob_rone
#               = (0.3*0.75) / (0.26)
```

Thus, the probability of a 1 transmitted given that a 1 was received is:

\[
P(T_1|R_1) = `r round(prob_tone_rone,4)`
\]