---
title: "Formative Assessment 4"
author: "Zyann Lynn Mayo"
date: "2025-03-02"
output: pdf_document
header-includes:
  - \usepackage{changepage}
  - \usepackage{tikz}
  - \usetikzlibrary{trees}

---
###### GitHub Link: https://github.com/mayonnayz/Probability-and-Probability-Distribution.git

## **Item 5: Law of Total Probability**

\begin{adjustwidth}{1cm}{1cm}
A geospatial analysis system has four sensors supplying images. The percentage of images supplied by each sensor and the percentage of images relevant to a query are shown in the following table.
\end{adjustwidth}

```{r}
sensor_data <- data.frame(
  Sensor = c(1, 2, 3, 4),
  Percentage_of_Images_Supplied = c(15, 20, 25, 40),
  Percentage_of_Relevant_Images = c(50, 60, 80, 85)
)

library(knitr)
kable(sensor_data, caption = "Sensor Data Table")

```

\begin{adjustwidth}{1cm}{1cm}
What is the overall percentage of relevant images?
\end{adjustwidth}

\bigskip

First, let's put the data into variables
\[
\begin{aligned}
P(S_1) &= 0.15, \quad P(R|S_1) = 0.50 \\ 
P(S_2) &= 0.20, \quad P(R|S_2) = 0.60 \\ 
P(S_3) &= 0.25, \quad P(R|S_3) = 0.80 \\ 
P(S_4) &= 0.40, \quad P(R|S_4) = 0.85
\end{aligned}
\]




```{r}
#Sensor 1
s1_img_sup <- .15
s1_rel_img <- .50
#Sensor 2
s2_img_sup <- .20
s2_rel_img <- .60
#Sensor 3
s3_img_sup <- .25
s3_rel_img <- .80
#Sensor 4
s4_img_sup <- .40
s4_rel_img <- .85
```

Using the law of total probability, for any event A:

\[
P(A) = \sum_{i=1}^{n} P(A \mid B_i) P(B_i)
\]

Inserting the given values,
```{r}
overall_rel_img <- (s1_img_sup*s1_rel_img) + 
  (s2_img_sup*s2_rel_img) + 
  (s3_img_sup*s3_rel_img) + 
  (s4_img_sup*s4_rel_img)
```

We get that the overall percentage of relevant images is **`r paste0(round(overall_rel_img * 100, 2), "%")`**


## **Item 6: Independence of Two Events**

\begin{adjustwidth}{1cm}{1cm}
A fair coin is tossed twice.
Let \( E_1 \) be the event that both tosses have the same outcome, that is,
\( E_1 = \{ HH, HT \} \). \\
Let \( E_2 \) be the event that the first toss is a head, that is, \( E_2 = \{ HH, HT \}\).\\
Let \( E_3 \) be the event that the second toss is a head, that is,
\( E_3 = \{ TH, HH \}\). \\
Show that \( E_1 \), \( E_2 \), and \( E_3 \) are pairwise independent but not mutually independent.
\end{adjustwidth}

Possible outcomes = {HH, HT, TH, TT}

\[
\begin{aligned}
E_1 = \{HH, TT\} = \frac{2}{4} = \frac{1}{2} \\
E_1 = \{HH, HT\} = \frac{2}{4} = \frac{1}{2} \\
E_1 = \{TH, HH\} = \frac{2}{4} = \frac{1}{2} 
\end{aligned}
\]

Insert values into variables:

```{r}
prob_e1 <- 1/2
prob_e2 <- 1/2
prob_e3 <- 1/2
```

#### Checking for Pairwise Independence \newline
 We have to check if \( P(E_i \cap E_j) = P(E_i)P(E_j) \) for all pairs. 

```{r}
library(MASS)
prob_e1_e2 <- fractions(prob_e1*prob_e2)
prob_e1_e3 <- fractions(prob_e1*prob_e3)
prob_e2_e3 <- fractions(prob_e2*prob_e3)
```
For \( E_1 \) and \( E_2 \): 

\[ 
\begin{aligned}
P(E_1 \cap E_2) = \{HH\} = \mathbf{1/4} \\ 
P(E_1)P(E_2) = 1/2 \times 1/2 = \mathbf{`r paste(prob_e1_e2)`} \\
P(E_1 \cap E_2) = P(E_1)P(E_2)
\end{aligned}
\] 
 
For \( E_1 \) and \( E_3 \): 

\[ 
\begin{aligned}
P(E_1 \cap E_3) = \{HH\} = \mathbf{1/4} \\ 
P(E_1)P(E_3) = 1/2 \times 1/2 = \mathbf{`r paste(prob_e1_e3)`}  \\
P(E_1 \cap E_3) = P(E_1)P(E_3)
\end{aligned}
\]

For \( E_2 \) and \( E_3 \): 

\[ 
\begin{aligned}
P(E_2 \cap E_3) = \{HH\} = \mathbf{1/4} \\ 
P(E_2)P(E_3) = 1/2 \times 1/2 = \mathbf{`r paste(prob_e2_e3)`}  \\
P(E_2 \cap E_3) = P(E_2)P(E_3)
\end{aligned}
\]

Thus, all pairs are **PAIRWISE INDEPENDENT**.

#### Checking for Mutual Independence \newline
 In order to prove that \( E_1, E_2, and E_3 \) are not mutually independent, we must show that
\[
P(E_1 \cap E_2 \cap E_3) \ne P(E_1)P(E_2)P(E_3)
\]

```{r}
library(MASS)
prob_e1_e2_e3 <- fractions(prob_e1*prob_e2*prob_e3)
```

\[
\begin{aligned}
P(E_1 \cap E_2 \cap E_3) = \{HH\} = \mathbf{1/4} \\
P(E_1)P(E_2)P(E_3) = 1/2 \times 1/2 \times 1/2 = \mathbf{`r paste(prob_e1_e2_e3)`}
\end{aligned}
\]

Since the probability of the intersection of the three events is 1/4, which is not equal to the product of the probability of all events which is 1/8, the events are **NOT MUTUALLY INDEPENDENT**


