---
title: "FA6 No 2"
author: "Zyann Lynn Mayo"
date: "2025-03-05"
output: pdf_document
---

# **II. Hypergeometric Distribution**

We need to find 
\[
P(X > 1) = 1 - P(X = 0) - P(X = 1)
\]
For both cases, we will follow the Hypergeometric Probability formula:
\[
P(X = k) = \frac{\binom{K}{k} \binom{N-K}{n-k}}{\binom{N}{n}}
\]

## Case 1: A sample of 10 is selected from a box of 40

Insert values into variables
```{r}
# k = selected number of defective chips
k1 <- 0.10
# n = sample size 
n1 <- 10
# N = population from given case
N1 <- 40
# K = defective 
K1 <- N1*k1

# Compute the probabilities P(X = 0) and P(X = 1)
P_X0_case1 <- dhyper(x = 0, m = K1, n = N1 - K1, k = n1)
P_X1_case1 <- dhyper(x = 1, m = K1, n = N1 - K1, k = n1)

# Compute P(X > 1) = 1 - P(X = 0) - P(X = 1)
P_greater1_case1 <- 1 - P_X0_case1 - P_X1_case1
```

We get that the probability that the sample contains more that 10% defectives when a sample of 10 is selected from a box of 40 is **`r paste0(round(P_greater1_case1 * 100, 2), "%")` **

## Case 2: A sample of 10 is selected from a box of 5000

Insert values into variables
```{r}
# k = selected number of defective chips
k2 <- 0.10
# n = sample size 
n2 <- 10
# N = population from given case
N2 <- 5000
# K = defective 
K2 <- N2*k2

# Compute the probabilities P(X = 0) and P(X = 1)
P_X0_case2 <- dhyper(x = 0, m = K2, n = N2 - K2, k = n2)
P_X1_case2 <- dhyper(x = 1, m = K2, n = N2 - K2, k = n2)

# Compute P(X > 1) = 1 - P(X = 0) - P(X = 1)
P_greater1_case2 <- 1 - P_X0_case2 - P_X1_case2
```

We get that the probability that the sample contains more that 10% defectives when a sample of 10 is selected from a box of 5000 is **`r paste0(round(P_greater1_case2 * 100, 2), "%")` **

