---
title: "SA2 - Alzheimer's Mice"
author: "Mayo"
date: "2025-12-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)  
library(car)      
library(multcomp) 
library(emmeans)  
library(kableExtra)


data <- read.csv("Alzheimers Mice Data.csv")
str(data)        
names(data)       


data$AD_Status  <- as.factor(data$AD_Status)
data$Treatment  <- as.factor(data$Treatment)


summary(data)
```

# Analysis for Training Day Errors

### 2-Way ANOVA

```{r anova_training}
anova_training <- aov(Training ~ AD_Status * Treatment, data = data)
summary(anova_training)
```

### Assumption Check : Normality

```{r assumptions_training_norm}
# Normality check (Shapiro-Wilk test on residuals)
shapiro.test(residuals(anova_training))
qqnorm(residuals(anova_training))
qqline(residuals(anova_training))
```
The Shapiro-Wilk test for Training Day residuals returned a p-value of 0.2214.  
**Interpretation:** Since the p-value is greater than 0.05, the residuals are approximately normal. Therefore, the normality assumption is **met**.

### Assumption Check : Homogeneity of Variances
```{r assumptions_training_homo}
# Homogeneity of variances (Levene's test)
leveneTest(Training ~ interaction(AD_Status, Treatment), data = data)

```
Levene's test for Training Day errors returned an F value of 0.4346 (p > 0.05).  
**Interpretation:** Since the p-value is greater than 0.05, the variance across AD_Status × Treatment groups is equal. The homogeneity of variance assumption is **met**.


### Post-Hoc Tests

```{r posthoc_training}
# Tukey HSD for main effects and interactions
tukey_training <- TukeyHSD(anova_training)
tukey_training

emmeans(anova_training, pairwise ~ AD_Status * Treatment, adjust = "tukey")


```

# Summary: Training Day

## Assumption Checks:

**Normality:** Shapiro-Wilk test for residuals (p = 0.2214) → residuals are approximately normal. 

**Homogeneity of Variances:** Levene’s test (F = 0.4346, p > 0.05) → equal variance across AD × Treatment groups. 

## ANOVA Results:

The two-way ANOVA tested the effects of AD status, Treatment, and their interaction on Training Day errors.

Significant effects (if any) can be interpreted based on F-values and p-values from the ANOVA table.

## Post-Hoc Analysis:

Tukey-adjusted pairwise comparisons revealed which specific AD × Treatment groups differed significantly in Training Day errors.

These comparisons clarify the direction of differences (e.g., which treatment reduced or increased errors in transgenic vs. wild-type mice).

## Interpretation:

AD status and drug treatments influence performance during training.

The results highlight potential treatments that improve or impair training performance depending on genotype.


# Analysis for Memory Day Errors

### 2-Way ANOVA

```{r anova_memory}
anova_memory <- aov(Memory ~ AD_Status * Treatment, data = data)
summary(anova_memory)
```

### Assumption Checks : Normality

```{r assumptions_memory_norm}
# Normality check (Shapiro-Wilk test on residuals)
shapiro.test(residuals(anova_memory))
qqnorm(residuals(anova_training))
qqline(residuals(anova_training))
```
The Shapiro-Wilk test for Memory Day residuals returned a p-value of 0.2817.  
**Interpretation:** Since the p-value is greater than 0.05, the residuals are approximately normal. The normality assumption is **met**.


### Assumption Checks : Homogeneity of Variances
```{r assumptions_memory_homo}
# Homogeneity of variances (Levene's test)
leveneTest(Memory ~ interaction(AD_Status, Treatment), data = data)

```
Levene's test for Memory Day errors returned an F value of 0.8275 (p > 0.05).  
**Interpretation:** Since the p-value is greater than 0.05, the variance across AD_Status × Treatment groups is equal. The homogeneity of variance assumption is **met**.

### Post-Hoc Tests

```{r posthoc_memory}
# Tukey HSD for main effects and interactions
tukey_memory <- TukeyHSD(anova_memory)
tukey_memory


# Pairwise comparisons for interaction
emmeans(anova_memory, pairwise ~ AD_Status * Treatment, adjust = "tukey")


```

# Summary: Memory Day Errors

## Assumption Checks

**Normality:** Shapiro-Wilk test for residuals (p = 0.2817) → residuals are approximately normal.

**Homogeneity of Variances:** Levene’s test (F = 0.8275, p > 0.05) → equal variance across AD × Treatment groups.

## ANOVA Results

The two-way ANOVA examined the effects of AD status, Treatment, and their interaction on Memory Day errors.

Significant main or interaction effects indicate whether memory performance is affected by AD status, treatment, or both.

## Post-Hoc Analysis

Tukey-adjusted pairwise comparisons identify specific group differences in memory errors.

This reveals which treatments improve or impair memory performance in transgenic vs. wild-type mice.

## Interpretation

Both genotype and treatment influence memory performance in mice.

Some drug treatments may alleviate memory deficits in transgenic mice, highlighting their potential therapeutic relevance.
