---
title: "SA2 No2"
author: "Zyann Lynn Mayo"
date: "2025-05-10"
output: 
  pdf_document:
    keep_tex: true
    latex_engine: pdflatex
---

## Test using Shapiro-Wilk normality test the Ethereum returns for trading data every five minutes, from August 7, 2015 to April 15, 2025.

NOTE: Due to source limitation, we were only able to retrieve the **DAILY** returns from **March 10, 2016** to **April 15, 2025**.

Data is retrieved from **investing.com** with this link.  \href{https://ph.investing.com/crypto/ethereum/historical-data?utm_source=google&utm_medium=cpc&utm_campaign=19321382621&utm_content=642152731664&utm_term=dsa-1638656206948_&GL_Ad_ID=642152731664&GL_Campaign_ID=19321382621&ISP=1&npl=1&af_adset_id=144619252957&ppu=9801673&gad_source=1&gad_campaignid=19321382621&gbraid=0AAAAAoxWJ7R5_S-EzRFNuwa7P7l-i6xI7&gclid=Cj0KCQjw8vvABhCcARIsAOCfwwqIw7tyASqJTQxG8zlPObz43VC3DoGpcIvtrEIKCommZT3xNt5uFDwaAqFTEALw_wcB}{(CLICK HERE)}

### First, we save the data from investing.com in a csv file. Then we retrieve the file.
```{r, message=FALSE}
suppressPackageStartupMessages(library(dplyr))
library(dplyr)
library(stats)

eth_data <- read.csv("Ethereum_Historical_Data.csv", stringsAsFactors = FALSE)

head(eth_data)

eth_data$Price <- as.numeric(gsub(",", "", eth_data$Price))

# Sort data by Date
eth_data <- eth_data %>%
  mutate(Date = as.Date(Date, format = "%b %d, %Y")) %>% 
  arrange(Date)

# Calculate log returns
eth_data <- eth_data %>%
  mutate(log_return = log(Price / lag(Price))) %>%
  filter(!is.na(log_return))

library(ggplot2)

ggplot(eth_data, aes(x = log_return)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 50,
                 fill = "skyblue",
                 color = "black",
                 alpha = 0.7) +
  stat_function(fun = dnorm,
                args = list(mean = mean(eth_data$log_return, na.rm = TRUE),
                            sd = sd(eth_data$log_return, na.rm = TRUE)),
                color = "red",
                linewidth = 1) +
  labs(title = "Histogram of Ethereum Daily Log Returns",
       x = "Log Return",
       y = "Density") +
  theme_minimal()


```

### Using the built in function, we run the Ethereum data through the Shapiro-Wilk test
```{r}
shapiro_test <- shapiro.test(eth_data$log_return)
```


After testing the normality of Ethereum daily log returns using the Shapiro-Wilk test. The result yielded a W statistic of **`r round(shapiro_test$statistic, 3)`** and a p-value of **`r ifelse(shapiro_test$p.value < 2.2e-16, "< 2.2e-16", round(shapiro_test$p.value, 3))`**.

Since the p-value is less than 0.05, the daily returns of Ethereum do **not** follow a normal distribution. This means the distribution of returns is significantly different from a normal distribution, possibly due to skewness, outliers, or fat tails. 

