---
title: "SA2 No2"
author: "Zyann Lynn Mayo"
date: "2025-05-10"
output: pdf_document
---

## Test using Shapiro-Wilk normality test the Ethereum returns for trading data every five minutes, from August 7, 2015 to April 15, 2025.

NOTE: Due to source limitation, we were only able to retrieve the **DAILY** returns from **March 10, 2016** to **April 15, 2025**.


### First, we retrieve the datas from the CSV file
```{r, message=FALSE}
suppressPackageStartupMessages(library(dplyr))
library(dplyr)
library(stats)

eth_data <- read.csv("Ethereum_Historical_Data.csv", stringsAsFactors = FALSE)

head(eth_data)

eth_data$Price <- as.numeric(gsub(",", "", eth_data$Price))

# Sort data by Date
eth_data <- eth_data %>%
  mutate(Date = as.Date(Date, format = "%b %d, %Y")) %>% 
  arrange(Date)

# Calculate log returns
eth_data <- eth_data %>%
  mutate(log_return = log(Price / lag(Price))) %>%
  filter(!is.na(log_return))

library(ggplot2)

ggplot(eth_data, aes(x = log_return)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 50,
                 fill = "skyblue",
                 color = "black",
                 alpha = 0.7) +
  stat_function(fun = dnorm,
                args = list(mean = mean(eth_data$log_return, na.rm = TRUE),
                            sd = sd(eth_data$log_return, na.rm = TRUE)),
                color = "red",
                linewidth = 1) +
  labs(title = "Histogram of Ethereum Daily Log Returns",
       x = "Log Return",
       y = "Density") +
  theme_minimal()


```

### Using the built in function, we run the Ethereum data through the Shapiro-Wilk test
```{r}
shapiro_test <- shapiro.test(eth_data$log_return)
```


After testing the normality of Ethereum daily log returns using the Shapiro-Wilk test. The result yielded a W statistic of **`r round(shapiro_test$statistic, 3)`** and a p-value of **`r ifelse(shapiro_test$p.value < 2.2e-16, "< 2.2e-16", round(shapiro_test$p.value, 3))`**.

Since the p-value is less than 0.05, the daily returns of Ethereum do **not** follow a normal distribution. This means the distribution of returns is significantly different from a normal distribution, possibly due to skewness, outliers, or fat tails. 

